---
title: "Final Project: Big Data in Immigration Research"
subtitle: "Our sexy title right here"
author: "Riya Chanduka, Dennis Hammerschmidt, Soyeon Jin, Pauline Kleinschlömer"
date: "`r Sys.Date()`"
output: 
  html_document:
  toc: true
---
## Introduction
Introduction 

 The number of migrants, especially forced migrants are increasing at a rapid phase worldwide. According to the 2018 report from the International Organization for Migration (IOM), 68.5 million individuals are forced to migrate in 2017 because of persecution, conflict, generalized violence, or human rights violations. The issue gets notably salient in 2015 following so-called ‘refugee crisis’ due to the Syrian civil war.
  Meanwhile, on 2nd September 2015, the picture of drowned 3-year-old Syrian boy, Aylan Kurdi in the Mediterranean Sea alarmed the world the seriousness of the issue and the difficulties which migrants face during their journey. However, despite its political saliency, there are still a lot of data gaps in migration research. IOM pointed out the issue in its 2018 report, insisting the necessity to fill out data gaps in the topics such as irregular migration, missing migrants and migration flows.
 “Missing Migrants Project” by International Organization for Migration (IOM) began to improve these situations after the tragic accident at October 2013, when two ships have wrecked near the Italian island of Lampedusa which caused the death of at least 368 individuals. Missing migrants project tracks the missing of refugees and asylum-seekers during their migration journey combining different data sources. By providing information about missing migrants, missing migrants project not only notify the world with the precarious situation migrants face but also point out the possible improvement in regulatory or administrative system to ensure safe migration journey. This achievement is important to tackle the United Nations’ Sustainable Development Goals (SDGs). Migration is a cross-cutting issue that is relevant to all SDGs. The Agenda’s core principle is “leaving no one behind”. Especially Goal 10.7 underlines an “orderly, safe, regular and responsible migration and mobility of people”. Therefore, it is mandatory to understand patterns of migrants’ journeys in order to achieve this goal by 2030.  The number of missing migrants can serve as a helpful indicator. However, the data of the project should be seen as approximations because the true number of missing migrants is unknown. It is not possible to report all deaths and disappearances because many bodies will never be found or identified.
 To dedicate to the missing migrants project aim, this report tries to explain the present of migration route and the dangerousness of different routes in the Mediterranean area, taking full advantage of the data offered by missing migrants project. Since migration is a continuous process that never stops, it is necessary to keep a constant eye on the changing patterns of migrants' travel. Therefore, this report analyses different routes of migrants in the Mediterranean Sea and compares changes over time. Also, this report tries to figure out possible factors affecting the journey to prevent further incident. To achieve this aim, this report begins by explaining the various migration routes, especially focusing on three different routes in the Mediterranean area. Then this report proceeds to explain the missing migrants project and the uniqueness of the data, along with the necessity to use big data for migration research. After that, death and missing of migrants on those routes are presented, suggesting the possible cause of dangerousness of the routes. Lastly, the report discusses the limitation of the report and suggest a possible improvement of the report. 


## Setup and basic data prepration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/dennis_hammerschmidt/Dropbox/missing_migrants/missing_migrants/")
```

```{r package_load, include=FALSE, cache=TRUE}
# Load all required packages and install those that you don't already have
install_and_load <- function(pkg) {
  new.pkgs <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if(length(new.pkgs))
    install.packages(new.pkgs, repos="https://cran.rstudio.com/")

  for(pkg_name in pkg)
    library(pkg_name,character.only=TRUE, quietly = TRUE)
}
packages <- c("ggplot2", "plotly", "readr", "RColorBrewer", "tidyverse", "grid", "dplyr", "gridExtra"
              ,"ggthemes", "ggfortify", "rworldmap", "lubridate", "reshape2", "forcats", "gganimate"
              ,"animation", "gifski", "epitools", "scales", "withr", "transformr", "devtools", "magrittr")
install_and_load(packages)
```

We're using the [Missing Migrants dataset](https://missingmigrants.iom.int/downloads) from 2014 to 2019 (up to the most recent dataset available). In our case, this is the one from May XX.

```{r data_and_preparation, include=FALSE, cache=TRUE, warning=FALSE}
# load the dataset
missing <- read_csv("MissingMigrants-Global-2019-05-09T10-44-32.csv")

# simplify column names
colnames(missing) <- c('id', 'region', 'date', 'year', 'month', 'num_dead', 'est_miss', 'total_dead_missing', 'num_surv', 'num_fem', 'num_male', 'num_child', 'cause', 'location', 'info', 'coord',
                      'route', 'url', 'geo_group', 'source_qual')

# create new data variable and fix the month variable
missing$month <- match(missing$month,month.abb)
missing$date2 <- paste0(missing$year, "/", missing$month, "/01")
date_helper <- as.month(missing$date2, format = "%Y/%m/%d")
missing$date2 <- date_helper$dates
missing$month <- format(missing$date2,"%m")

# create longitude and latitude from the coordinates variable and make them numeric
missing <- separate(missing, into = c("lat","lon"), coord, sep = ",")
missing$lon <- as.numeric(missing$lon)
missing$lat <- as.numeric(missing$lat)

# filter for missings in the date, latitude and longitude
missing <- missing %>% 
  filter(!is.na(date2) & !is.na(lon) & !is.na(lat))

# replace NAs in missing and dead with 0 (for computational purposes)
missing[c("est_miss", "num_dead", "num_surv",
          "num_fem", "num_male", "num_child")][is.na(missing[c("est_miss", "num_dead", "num_surv",
          "num_fem", "num_male", "num_child")])] <- 0

# replace NAs in route and region with unkown
missing[c("route")][is.na(missing[c("route")])] <- "Unkown"
missing[c("region")][is.na(missing[c("region")])] <- "Unkown"

# subset the dataset to include only variables of interest

missing <- subset(missing, select = c("id", "region", "date2", "year", "month", "num_dead", "est_miss", "total_dead_missing", "num_surv", "num_fem", "num_male", "num_child", "cause", "lon", "lat", "route", "source_qual"))
```

## First descriptive: Where did people go missing the most?

```{r regions_and_year, warning=FALSE}
# create regions data that includes the sum of missings and dead migrants per year for each region
regions <- missing %>%
  group_by(region,year) %>%
  dplyr::summarise(sum(est_miss), sum(num_dead))

regions <- as.data.frame(regions)
colnames(regions) <- c("region","year","missing","dead")
regions <- melt(regions, id.vars = c("region", "year"))

# define position ordering for graph
positions <- c("Unknown", "East Asia", "North America", "South America", "Caribbean", "Europe", "Sub-Saharan Africa", "Central America incl. Mexico", "Horn of Africa", "Southeast Asia", "U.S./Mexico Border", "North Africa", "Mediterranean")

# plot the total number of missings and dead across regions and years
incident_region <- ggplot(regions) +
  geom_bar(aes(x = region, y = value, fill = variable),
           stat = 'identity',
           position = 'dodge') +
  coord_flip() +
  labs(x = "Regions", y = "Total number of missings/deaths") +
  scale_x_discrete(limits = positions) +
  labs(fill = "Incident") +
  facet_wrap(vars(year))

# use ggplotly for interactive exploration
ggplotly(incident_region)
```

As we see and suspected before, most incidences happen at the Mediterranean Sea.

```{r stacked_bar_total, message=FALSE, warning=FALSE}
# subset only for the Mediterranean Sea
missing_medsea <- subset(missing, region=="Mediterranean")

# create regions data only for Mediterranean Sea that also includes the number of people surviving an incident
region_med <- missing_medsea %>%
  group_by(route,year) %>%
  dplyr::summarise(sum(est_miss), sum(num_dead), sum(num_surv))

region_med <- as.data.frame(region_med)
colnames(region_med) <- c("route","year","missing","dead", "survived")
region_med <- melt(region_med, id.vars = c("route", "year"))

# reduce the name of the routes to their direction (since we're only in the Mediterranean Sea)
route_short <- c("Central Mediterranean"="Central" ,"Western Mediterranean"="Western", "Eastern Mediterranean"="Eastern")
region_med$route <- as.character(route_short[region_med$route])

# define position ordering for graph
positions2 <- c("Central", "Western", "Eastern")

# plot the total number of missings and dead across regions and years
surv_or_not <- ggplot(region_med) +
  geom_bar(aes(x = route, y = value, fill = variable), 
               stat = 'identity', position = 'fill') +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Regions", y = "Percentage of incidents") +
  scale_x_discrete(limits = positions2) +
  labs(fill = "Incident") +
  facet_wrap(vars(year)) 

# use ggplotly for visualization and round the percentage of stacked bar chart by 2 digits
with_options(list(digits = 1), ggplotly(surv_or_not)) 
```

```{r map_static, cache=TRUE, message=FALSE, warning=FALSE}
# intiatlize a simple map
worldMap <- fortify(map_data("world"), region = "mediterranean")

# define axis, shape and visual display of the map
map <-ggplot() + 
  geom_map(data = worldMap, map = worldMap,aes(x = long, y = lat, map_id = region, group = group),fill = "white", color = "black", size = 0.25)

# plot the map with total number of dead and missing migrants including the three routes in the Mediterranean Sea
med_map <- map + geom_point(aes(x = lon, y = lat, size=total_dead_missing), alpha=0.5, data = missing_medsea) +
        geom_path(aes(lon,lat,col=route),data=missing_medsea,alpha=0.2) +
        theme_fivethirtyeight(base_size = 10, base_family = "sans") + 
        scale_color_brewer(palette='Set1') + 
        theme(axis.text.x = element_text(size=8,angle=90),legend.position='top') +
        xlab('') + ylab('') + 
        ggtitle('Total incidents in the Mediterranean Sea') + ylim(c(29, 47)) + 
      xlim(c(-8, 42))
med_map
# optionally: use ggplotly for visualization. Caution: takes some time to run!
#ggplotly(med_map)
```

```{r map_animated, cache = TRUE, message=FALSE, warning=FALSE}
# If not available, load the transformr package directly for github using the following command(s):
# devtools::install_github("thomasp85/transformr") # load if needed
# library(transformr)

# create an animated map for the total number of dead and missing migrants as well as the three routes on the Mediterranean Sea over the entire time frame of the dataset
mapanimated <- map + geom_point(aes(x = lon, y = lat, size=total_dead_missing), alpha=0.5, data = missing_medsea) +
      geom_path(aes(lon,lat,col=route),data=missing_medsea,alpha=0.3) +
      theme_fivethirtyeight(base_size = 10, base_family = "sans") + 
      scale_color_brewer(palette='Set1') + 
      theme(axis.text.x = element_text(size=8,angle=90),legend.position='top') +
      xlab('') + ylab('') + 
      ggtitle('Total incidents in the Mediterranean Sea') + ylim(c(29, 47)) + 
  xlim(c(-8, 42)) +
  labs(title = 'Year: {frame_time}') +
  transition_time(year) 

mapanimated

# optional: save the animation as .gif
#anim_save("mapanimated.gif", animation = last_animation())

```


